Automatically proving equivalence by type-safe reflection

### 2016-01-13 paper-JFP-DTP/content/2-basicIdeas.tex
Added the source file for the example with Nat

\x => \y => \z => Just (replace (sym (replace (sym (replace 
(plusZeroRightNeutral x) (plusAssociative x 0 y))) (replace 
(sym (replace (plusZeroRightNeutral x) (plusAssociative x 0 z))) (replace 
(replace (plusZeroRightNeutral (plus xy)) (plusAssociative (plus x y) 0 x)) 
[...]
: (x : Nat) -> (y : Nat) -> (z : Nat) 
  -> Maybe ((x + y) + x + z 
            = x+ (y + x) + z)
\end{verbatim}
\end{center}
\caption{Obtained proof}
\figrule
\end{figure}

And we effectively get the proof of equality we wanted. As expected, this proof uses the properties of associativity ($plusAssociative$) and the property of neutrality of $Z$ for $plus$ ($plusZeroRightNeutral$).

@@

We still want an automatic way of constructing these reflected terms because what we have so far is enough to demonstrate that our "prover" works, but that's not enough for being used intensively. 
By using Idris's reflection mechanism, we can tag a function with the keyword "\%reflection", which means that this function runs on syntax instead of values. That implies that when applying this function to some arguments, the arguments aren't reduce before the evaluation of this function.

>>>>

\x => \y => \z => Just (replace (sym (replace (sym (replace 
(sym (plusAssociative x 0 y)) (replace (replace (sym 
(plusZeroRightNeutral x)) Refl) Refl))) (replace (sym 
(replace (sym (plusAssociative x 0 z)) (replace (replace 
(sym (plusZeroRightNeutral x)) Refl) Refl))) (replace (sym 
(plusAssociative (x+y) x z)) [...]
: (x : Nat) -> (y : Nat) -> (z : Nat) 
  -> Maybe ((x + y) + x + z 
            = x + (y + x) + z)
\end{verbatim}
\end{center}
\caption{Obtained proof}
\figrule
\end{figure}

And we effectively get the proof of equality we wanted. As expected, this proof uses the properties of associativity ($plusAssociative$) and the property of neutrality of $Z$ for $+$ ($plusZeroRightNeutral$).

@@

We still want an automatic way of constructing these reflected terms because what we have so far is enough to demonstrate that our "prover" works, but that's not enough for being used intensively. We want to program an automatic way of going from the concrete values (of type $Nat$), to the reflected terms (of type $Expr$). The only way to do that is to inspect the abstract syntax tree of the concrete value.
By using Idris's reflection mechanism, we can tag a function with the keyword "\%reflection", which means that this function runs on syntax instead of values. That implies that when applying this function to some arguments, the arguments aren't reduce before the evaluation of this function.
